% Discussion
\subsection{Maximum Intensity or Mean Intensity?}

\subsection{Future Directions}
A future direction of this work to to move toward fully automated segmentation and calculation vessel diameters from enface angiogram maps.  While ROIs were manually selected in this work, it is also possible to automatically segment angiographic data using traditional skeletonization methods, which are commonly used for automated OCTA analysis.  From a skeletonized image, vessel angle can be computed for further extraction of vascular cross-sections perpendicular to the vessel axis.  These cross-sections could then be processed using the model-based methods described here to enable fully automated analysis of 3D OCTA data sets, rather than the semi-automated approach shown here for 2D OCTA data.  This type of model-based approach is particularly important for comparing 3D volumes acquired at different time points in the mouse eye as there are often intensity losses caused by drying of the eye, cataract formation, etc.  As we have shown in this work, even low levels of intensity loss occurring over a short time window can significantly impact the measured vessel diameter when using a binarization approach, but not when using the models presented here.

\subsection{Computation Time} 
It is clear from the results that model-based approaches outperform traditional diameter measurements based on a binarized image.  Nevertheless, the improved linearity and repeatability come at the cost of computation time.  Fitting a model to the raw aniogram data is significantly more computationally expensive and may not always be the most appropriate choice if results of the processing are time sensitive.  Nevertheless, these approaches are highly parallelizable given the independent nature of the measurements, so it is expected that compuation time could be significantly accelerated with appropriate hardware.

\subsection{Threshholding}
The results presented here used a basic, global threshold for the binarization of the angiogram data.  There are numerous other approaches for binarizing the angiogram, which will most likely each behave differently.  For example, angiogram signals can be enhanced with vesselness filters and dynamic thresholds can be used to binarize images with varying SNR.  When applying a vesselness filter, it is possible to enhance the shape of the angiogram and omit noise, but it is also possible for such enhancement to overcompensate the signal, leading to larger than expected vessel appearance.  Dynamic thresholding can also yield a more uniform angiogram appearance, but similarly risks over- or undercompensating vessels given that threshold level directly affects the measured diameter.  Further analysis of these other methods would be required to determine their efficacy in returning accurate and precise diameter values. 