% Methods and Materials
%\section{Methods and materials}

\subsection{Imaging system}
\label{subsec:imaging_system}
All imaging was performed using a custom-built polarisation-sensitive spectral-domain OCT ophthalmoscope, described in detail previously~\cite{fialovaPolarizationPropertiesSingle2016}. The system was centred at 840~nm with a spectral bandwidth of 100~nm at full width half maximum (FWHM), providing an axial resolution of approximately 5.1 µm in air and 3.8 µm in tissue (assuming a refractive index of 1.35). A sensitivity of 96 dB was achieved at A-scan rate of 80~kHz and a beam diameter of 0.5~mm at the pupil plane (power at cornea~=~2.85~mW)~\cite{merkleHighresolutionDepthresolvedVascular2021}. This system was employed to acquire volumetric OCT data from mouse retinas before, during, and after an intravascular contrast agent administration, as well as before and after functional stimulation of the mouse retina using a 502~nm light-emitting diode (LED) ring flashing at 10~Hz (power at cornea~$= 26.6\,\mu\mathrm{W}$). This wavelength was selected to target the peak absorption of mouse rhodopsin ($\sim$500~nm), enabling predominantly rod-driven photoreceptor stimulation~\cite{Son2017,sonFunctionalOCTReveals2024}.

\subsection{Experimental protocol}
\label{subsec:experimental_protocol}
Three mouse models were used in this study: VLDLR knockout mice (B6;129S7-\emph{Vldlr\textsuperscript{tm1Her}}/J, The Jackson Laboratory, Bar Harbor, USA), C57BL/6 mice \textcolor{red}{(C57BL/6, The Jackson Laboratory, Bar Harbor, USA)}, and SOD1 transgenic mice \textcolor{red}{(strain details, The Jackson Laboratory, Bar Harbor, USA)}, \textcolor{red}{with $n$~=~XX mice per group}. All imaging sessions were conducted under general anaesthesia using either isoflurane or a ketamine/xylazine combination. Isoflurane was induced at 4\% in oxygen for 4 minutes and subsequently maintained at 2\% in oxygen at a flow rate of 1.5--2~L/min. In cases where isoflurane alone did not provide adequate sedation, an intraperitoneal injection of ketamine/xylazine cocktail (100~mg/kg ketamine, 6~mg/kg xylazine; 10~mL/kg body weight) was administered. To maintain core body temperature of the mice throughout imaging, they were positioned on a heating pad. Mice pupils were dilated with one drop of tropicamide (5~mg/mL, Agepha Pharma s.r.o., Senec, Slovakia) per eye, and artificial tear drops (Oculotect, ALCON Pharma GmbH, Freiburg im Breisgau, Germany) were applied at regular intervals to prevent cornea from drying out.

Intralipid has been shown to be a highly scattering contrast agent that fills the vascular lumen uniformly, thereby compensating for the artefact caused by the diminished backscattering at vessel peripheries caused by RBC orientation effects~\cite{merkleHighresolutionDepthresolvedVascular2021,panUltrasensitiveDetection3D2014,cimallaShearFlowinducedOptical2011}. For this study, Intralipid 20\% (3~mL/kg body weight) was administered as a bolus injection via the mice tail vein. Volumetric angiography data were acquired over a 1~mm~$\times$~1~mm field of view centred on the optic nerve head (ONH), with a scan density of 500 A-scans per B-scan and five repeated B-scans at each of 400 slow-axis positions (acquisition time approximately 16~seconds). For this study, pre- and post-injection datasets and pre- and post-stimulation datasets were acquired. Additionally, a dynamic-contrast OCT (DyC-OCT) protocol~\cite{merkleLaminarMicrovascularTransit2016} was employed pre- and post-contrast injection, consisting of 2000 repeated B-scans at the same cross-sectional location (interscan time 8.2~ms) over a 1~mm lateral field of view.~\textcolor{red}{All animal procedures were approved by the ethics committee of the Medical University of Vienna and the Austrian Federal Ministry of Education, Science, and Research98 (BMBWF/66.009/0272-V/3b/2019)}.

\subsection{Post-processing}
\label{subsec:post_processing}
Cross-polarised OCT intensity data from the retinal pigment epithelium (RPE) was used for inter-frame alignment and B-scan flattening~\cite{augustinMultifunctionalOCTImage2016,merkleHighresolutionDepthresolvedVascular2021}. Angiograms were derived by applying the complex subtraction OCT angiogram (OCTA) algorithm~\cite{srinivasanRapidVolumetricAngiography2010,harperRetinalAnalysisMouse2020}. To enhance vessel visibility and suppress background noise, a 3D averaging filter with a kernel size of 50~$\times$~50~$\times$~10 voxels was applied to both the angiogram and DyC-OCT volumes. For the volumetric angiogram data, the kernel dimensions correspond to the lateral fast-axis ($x$), lateral slow-axis ($y$), and axial ($z$) dimensions, while for the DyC-OCT data, the kernel spans the lateral ($x$), axial ($z$), and temporal ($t$) dimensions. All data processing and analysis were performed in MATLAB (R2019b, MathWorks, Natick, MA, USA).

\subsection{Diameter measurement}
\label{subsec:diameter_measurement}
All vessel diameter measurements were performed on flattened B-scan angiograms. In this study, we compare three projection-based approaches and two model-based approaches for measuring vessel diameter from OCTA data, using post-contrast measurements as a ground truth reference. All measurements were performed in linear intensity scale. Vessel intensity profiles were measured from en-face intensity profiles generated by projecting the A-scans within a rectangular region of interest (ROI) encompassing the vessel cross-section along the lateral ($x$) and axial ($z$) directions. Vessel diameter was then calculated by fitting parametric models to these one-dimensional intensity profiles, extracted from the B-scan data across each vessel of interest. The details about the same are described in the following sections.

\subsubsection{Ground truth calculation}
\label{subsubsec:ground_truth}
Ground truth vessel diameters were obtained from contrast-enhanced B-scan angiograms, where the uniformly scattering Intralipid fills the full vascular lumen regardless of RBC orientation. For each vessel of interest, an elliptical ROI was manually selected around the vessel cross-section in the flattened B-scan. The lateral axis of the ellipse (perpendicular to the OCT beam) was taken as the ground truth vessel diameter, as this corresponds to the dimension represented in en-face OCTA projection maps. In cases where the lateral axis length exceeded the axial axis length (i.e.~when the lateral dimension was the major axis of the ellipse), eccentricity correction was performed by correcting the measured lateral length by multiplying with the ratio of the minor to the major axis length. This correction compensated for the apparent broadening of the vessel cross-section caused by an oblique intersection of the imaging plane with the vessel axis.

\subsubsection{Vessel projection calculation}
\label{subsubsec:vessel_projection}
For each vessel of interest, a rectangular ROI encompassing the vessel cross-section was selected from the DyC-OCT B-scan angiograms. En-face intensity profiles were then generated by projecting the A-scans within this ROI along the lateral ($x$) and axial ($z$) directions. Two projection methods were compared: maximum intensity projection (MaIP), which selects the highest intensity value along each A-scan, and mean intensity projection (MeIP), which averages the values along each A-scan. The resulting one-dimensional profiles were normalised to a range of 0--1 to ensure uniformity across vessels, scans, and animals. Additionally, for comparison with conventional OCTA processing pipelines that rely on image binarisation, a binary mask was generated by applying a global threshold to the en-face projection maps. These three approaches: mean projection, maximum projection, and binary thresholding, represent commonly used methods in the literature for OCTA quantification~\cite{untrachtTowardsStandardisingRetinal2024,hormelMaximumValueProjection2018,sampsonTowardsStandardizingRetinal2022,leeGeneratingRetinalFlow2019}.

\subsubsection{Vessel diameter calculation}
\label{subsubsec:vessel_dameter}
Two parametric models were then fitted to these one-dimensional intensity profiles from the previous step, using nonlinear least-squares regression in MATLAB.

The first model was a standard Gaussian function:
\begin{equation}
\label{eq:gaussian}
f(x) = A \exp\!\left(-\frac{{(x - \mu)}^2}{2\sigma^2}\right) + C
\end{equation}
where $A$ is the peak amplitude, $\mu$ is the centre position, $\sigma$ is the standard deviation, and $C$ is a constant baseline offset.

The second model was a generalised Gaussian (GG) function:
\begin{equation}
\label{eq:gen_gaussian}
f(x) = A \exp\!\left(-{\left(\frac{|x - \mu|}{\alpha}\right)}^{\!\beta}\right) + C
\end{equation}
where $A$ is the peak amplitude, $\mu$ is the centre position, $\alpha$ is the scale parameter, $\beta$ is the shape parameter, and $C$ is a constant baseline offset. The shape parameter $\beta$ was constrained to $\beta \geq 2$ during fitting. Given the imaging system configuration, all vessels are expected to produce at minimum a Gaussian intensity profile in the B-scan cross-section, and so profiles sharper than Gaussian profile ($\beta < 2$) are therefore not technically meaningful. With this constraint, the GG accommodates profiles ranging from Gaussian ($\beta = 2$, where $\sigma = \alpha / \sqrt{2}$) to increasingly flat-topped profiles ($\beta > 2$).

Two standard width metrics were calculated from the fitted models: the full width at half maximum (FWHM) and the $1/e^2$ width. For the standard Gaussian fit, the $1/e^2$ width was calculated as:
\begin{equation}
\label{eq:gauss_1e2}
w_{1/e^2} = 4\sigma
\end{equation}
The FWHM for a Gaussian is related to the $1/e^2$ width by a fixed proportionality:
\begin{equation}
\label{eq:gauss_fwhm}
\mathrm{FWHM} = 2\sigma\sqrt{2\ln 2} = \frac{\sqrt{2\ln 2}}{2} \; w_{1/e^2}
\end{equation}
and therefore only the $1/e^2$ width was calculated directly from the fit, with the FWHM obtainable via this known relationship. 

For the GG fit, no such linear relationship exists between the FWHM and $1/e^2$ width, since both depend on the shape parameter $\beta$. Both metrics were therefore calculated independently as:
\begin{equation}
\label{eq:gg_fwhm}
\mathrm{FWHM}_{\mathrm{GG}} = 2\alpha{\left(\ln 2\right)}^{1/\beta}
\end{equation}
\begin{equation}
\label{eq:gg_1e2}
w_{1/e^2,\,\mathrm{GG}} = 2\alpha \cdot 2^{1/\beta}
\end{equation}

For the binary mask approach, multiple threshold values were used, and the resulting vessel diameters were compared to the ground truth. For each C-scan, the maximum angiogram intensity within the smoothed volume was calculated, and the threshold values were a fixed percentage of this maximum intensity. The vessel diameter was then measured as the number of pixels above each of the threshold values across the vessel cross-section.